19Q2

### 2019-06-04 19Q2.Rmd

# Manually assign biomarker_types based on a combination of biomarker_type derived from the JSON string and patterns in the Feature.Names
biomarker_type_curated <- 
  ifelse(!is.na(str_match(string = tolower(g2p$biomarker_type), pattern = "copy number variant|amplification")) | !is.na(str_match(string = tolower(g2p$Feature.Names), pattern = "copy number variant|amplification")), "Amplification/CNV",
  ifelse(!is.na(str_match(string = toupper(g2p$Feature.Names), pattern = "[A-Z0-9|A-Z]+-[A-Z0-9|A-Z]+|FUSION")) & is.na(str_match(string = g2p$Feature.Names, pattern = "[0-9]+-[0-9]+")) | g2p$biomarker_type == "fusion" | g2p$biomarker_type == "rearrange", "Fusion",
  ifelse(!is.na(str_match(string = g2p$Feature.Names, pattern = "[A-Z][0-9]+([A-Z]|[*])|[0-9A-Z]+ [A-Z][0-9]+|[A-Z][0-9]+$")), "Mutation",
  ifelse(!is.na(str_match(string = tolower(g2p$Feature.Names), pattern = "expression|methylation")), "Expression",
  ifelse(!is.na(str_match(string = g2p$biomarker_type, pattern = "intron|insertion|deletion|mutant")), "Mutation",
  ifelse(!is.na(str_match(string = tolower(g2p$Feature.Names), pattern = "codon|positive|biallelic inactivation|del|ins|dup|mut|frameshift|missense|variant|[0-9]+fs|>")), "Mutation",
  ifelse(!is.na(str_match(string = g2p$Feature.Names, pattern = "^KIT$|^BRCA$")), "Mutation", NA)))))))
g2p$biomarker_type_curated <- biomarker_type_curated[,1]

# Edit direction column
g2p <- add_column(g2p, Direction_Curated = NA, .after = "Direction")
g2p$Direction_Curated <- ifelse(!is.na(str_match(string = tolower(g2p$Direction), pattern = "resist")), "resistance", ifelse(!is.na(str_match(string = tolower(g2p$Direction), pattern = "sensitive|sensitivity|responsive")), "sensitivity", "none"))

@@

Mutation calls (coding region, germline filtered):
```{r mutation_status, eval=FALSE}
maf_raw <- read.delim("./data_munging/19Q2/CCLE_mutations_20190514.csv.gz", header = TRUE, sep = ",")
maf_raw[, 1:2] <- NULL
colnames(maf_raw)[colnames(maf_raw) == "Hugo_Symbol"] <- "Gene"
maf_raw <- unfactorize(maf_raw)
# Select columns
maf <- subset(maf_raw, select = c("Gene", "DepMap_ID"))
maf$Mutation <- 1
maf_grid <- expand.grid("Gene" = unique(maf$Gene), "DepMap_ID" = unique(maf$DepMap_ID))
maf_df <- merge(maf, maf_grid, by = c("Gene", "DepMap_ID"), all = TRUE)
maf_df[is.na(maf_df)] <- 0

saveRDS(maf_df, "./data_munging/19Q2/maf_df_19Q2.rds", compress = "xz")
```

```{r}
cgi <- read.delim("./../../../Desktop/cgi_biomarkers_latest/cgi_biomarkers_per_variant.tsv", sep = "\t", header = TRUE)
```

@@

```{r omics_functions}
MergeOmicsG2P <- function(database, maf_df, ge_df, cn_df, gene_list, outfile) {
  # Merge mutation status, CN, and GE omics data with drugscreen results and metadata
  if(database %in% c("CCLE", "CTRP", "GDSC")) {
    filt <- auc %>% filter(Database == database) %>% select(Cell_line, CCLE_histology, TCGA_classification, CCLE_name, DepMap_ID, Drug, AUC_IC50)
    print(paste0("Processing ", database, " data."))
    colnames(filt)[colnames(filt) == "AUC_IC50"] <- "Score"

    grid <- expand.grid("Cell_line" = unique(filt$Cell_line), "Gene" = gene_list)
    data <-  merge(filt, grid, by = "Cell_line", all.y = TRUE)
    print("Merged grid.")
    # Merge metadata
    data <-  merge(data, select(ccl_meta, Cell_line, CCLE_histology, CCLE_name, DepMap_ID, disease), by = c("Cell_line", "CCLE_histology", "CCLE_name", "DepMap_ID"), all.y = TRUE)
    ## Add columns for G2P drug-gene and drug-gene-lineage sorting
    data$Drug_Gene <- paste0(data$Drug, "_", data$Gene)
    data$G2P_Drug_Gene <- ifelse(data$Drug_Gene %in% g2p_dg, "Yes", "No")
    data$Drug_Gene_Lin <- paste0(data$Drug, "_", data$Gene, "_", data$disease)
    data$G2P_Drug_Gene_Lin <- ifelse(data$Drug_Gene_Lin %in% g2p_dgl, "Yes", "No")
    data$Gene_Lin <- paste0(data$Gene, "_", data$Gene, "_", data$disease)
    data$G2P_Gene_Lin <- ifelse(data$Gene_Lin %in% g2p_gl, "Yes", "No")
    print("Merged metadata.")
    # Merge mutation calls
    data <- unfactorize(data)
    data <- merge(data, maf_df, by = c("Gene", "DepMap_ID"), all.x = TRUE)
    data$Mutation <- ifelse(is.na(data$Mutation), 0, data$Mutation)
  }
  if(database == "CRISPR") {
    print(paste0("Processing ", database, " data."))
    # Melt CRISPR dataset for merging
    data <- melt(crispr_all, id.vars = "DepMap_ID", measure.vars = colnames(crispr_all)[2:ncol(crispr_all)], variable.name = "Gene", value.name = "Score")
    # Merge cell line metadata
    data <- merge(data, select(ccl_meta, Cell_line, CCLE_histology, CCLE_name, DepMap_ID, disease), by = "DepMap_ID", all.y = TRUE)
    print("Merged metadata.")
    # Merge mutation calls
    data <- unfactorize(data)
    data <- merge(data, maf_df, by = c("Gene", "DepMap_ID"), all.x = TRUE)
    data$Mutation <- ifelse(is.na(data$Mutation), 0, data$Mutation)
  }

@@

`r nrow(filter(test, n != 1))`/`r nrow(test)` gene-cell line combinations have more than one mutation in the MAF file. This is corrected in the point mutation filtering done below.
```{r merge_drug_data, eval=FALSE}
# Load in Pozdeyev et al. (2016) data
ccle_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CCLE_drug_response_metrics.csv.gz", sep = ",")
ctrp_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CTRP_drug_response_metrics.csv.gz", sep = ",")
gdsc_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/GDSC_drug_response_metrics.csv.gz", sep = ",")

auc <- bind_rows(ccle_all, ctrp_all, gdsc_all)

# Harmonize Pozdeyev cell line names with DepMap
auc$Drug <- tolower(auc$Drug_name)
auc$Cell_line <- toupper(gsub("-", "", auc$Cell_line))
auc$Cell_line <- ifelse(auc$Cell_line == "K052", "KO52", auc$Cell_line)
auc$Cell_line <- ifelse(auc$Cell_line == "OVCAR3", "NIHOVCAR3", auc$Cell_line)
auc$Cell_line <- ifelse(auc$Cell_line == "OVCAR3", "NIHOVCAR3", auc$Cell_line)
auc$Cell_line <- ifelse(auc$Cell_line == "TI73", "T173", auc$Cell_line)
auc <- merge(ccl_histologies, auc, by = "Cell_line", all.y = TRUE)

ccle <- MergeOmicsG2P(database = "CCLE", maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = g2p_genes, outfile = "./data_munging/19Q2/ccle_data_19Q2.rds")
write.csv.gz(ccle %>% drop_na(Score), "./data_munging/19Q2/ccle_data_19Q2.csv.gz")

ctrp <- MergeOmicsG2P(database = "CTRP", maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = g2p_genes, outfile = "./../htscreens_giant_files/ctrp_data_19Q2.rds")
write.csv.gz(ctrp %>% drop_na(Score), "./../htscreens_giant_files/ctrp_data_19Q2.csv.gz")

gdsc <- MergeOmicsG2P(database = "GDSC", maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = g2p_genes, outfile = "./data_munging/19Q2/gdsc_data_19Q2.rds")
write.csv.gz(gdsc %>% drop_na(Score), "./../htscreens_giant_files/gdsc_data_19Q2.csv.gz")

# all_genes <- intersect(unique(maf_df$Gene), intersect(unique(cn_melt$Gene), unique(ge_melt$Gene)))
# ccle <- MergeOmicsG2P(database = "CCLE", maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = all_genes, outfile = "./../htscreens_giant_files/ccle_data_19Q2.rds")
# ctrp <- MergeOmicsG2P(database = "CTRP", maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = all_genes, outfile = "./../htscreens_giant_files/ctrp_data_19Q2.rds")
# gdsc <- MergeOmicsG2P(database = "GDSC", maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = all_genes, outfile = "./../htscreens_giant_files/gdsc_data_19Q2.rds")
```

@@

Merge fusions:
```{r, eval=FALSE}
ccle_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CCLE_drug_response_metrics.csv.gz", sep = ",")
ctrp_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CTRP_drug_response_metrics.csv.gz", sep = ",")
gdsc_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/GDSC_drug_response_metrics.csv.gz", sep = ",")

auc <- bind_rows(ccle_all, ctrp_all, gdsc_all) %>% drop_na(Drug_name)

# Harmonize Pozdeyev cell line names with DepMap
auc$Drug <- tolower(auc$Drug_name)
auc$Cell_line <- toupper(gsub("-", "", auc$Cell_line))
auc$Cell_line <- ifelse(auc$Cell_line == "K052", "KO52", auc$Cell_line)
auc$Cell_line <- ifelse(auc$Cell_line == "OVCAR3", "NIHOVCAR3", auc$Cell_line)
auc$Cell_line <- ifelse(auc$Cell_line == "OVCAR3", "NIHOVCAR3", auc$Cell_line)
auc$Cell_line <- ifelse(auc$Cell_line == "TI73", "T173", auc$Cell_line)
auc <- merge(ccl_histologies, auc, by = "Cell_line", all.y = TRUE)

fusions_filt <- filter(fusions, FusionPair %in% unique(as.character(g2p_fus$FusionPair)))

ccle_fus <- auc %>% filter(Database == "CCLE") %>% select(Cell_line, CCLE_histology, TCGA_classification, CCLE_name, DepMap_ID, Drug, AUC_IC50)
ccle_fus_grid <- expand.grid("DepMap_ID" = unique(ccle_fus$DepMap_ID), "FusionPair" = unique(fusions_filt$FusionPair))
ccle_fus <- merge(merge(ccle_fus, ccle_fus_grid, by = "DepMap_ID", all = TRUE), fusions_filt, by = c("DepMap_ID", "CCLE_name", "CCLE_histology", "Cell_line", "FusionPair"), all = TRUE)
ccle_fus$Fusion_Status <- ifelse(!is.na(ccle_fus$LeftGene), "Yes", "No")
ccle_fus$Drug_FusionPair <- ifelse(is.na(ccle_fus$Drug) | is.na(ccle_fus$FusionPair), NA, paste0(ccle_fus$Drug, "_", ccle_fus$FusionPair))
ccle_fus$G2P_Drug_FusionPair <- ifelse(ccle_fus$Drug_FusionPair %in% unique(as.character(g2p_fus$Drug_FusionPair)), "Yes", "No")
ccle_fus$G2P_Fusion <- ifelse(ccle_fus$FusionPair %in% unique(as.character(g2p_fus$FusionPair)), "Yes", "No")
ccle_fus$Database <- "CCLE"

ctrp_fus <- auc %>% filter(Database == "CTRP") %>% select(Cell_line, CCLE_histology, TCGA_classification, CCLE_name, DepMap_ID, Drug, AUC_IC50)
ctrp_fus_grid <- expand.grid("DepMap_ID" = unique(ctrp_fus$DepMap_ID), "FusionPair" = unique(fusions_filt$FusionPair))
ctrp_fus <- merge(merge(ctrp_fus, ctrp_fus_grid, by = "DepMap_ID", all = TRUE), fusions_filt, by = c("DepMap_ID", "CCLE_name", "CCLE_histology", "Cell_line", "FusionPair"), all = TRUE)
ctrp_fus$Fusion_Status <- ifelse(!is.na(ctrp_fus$LeftGene), "Yes", "No")
ctrp_fus$Drug_FusionPair <- ifelse(is.na(ctrp_fus$Drug) | is.na(ctrp_fus$FusionPair), NA, paste0(ctrp_fus$Drug, "_", ctrp_fus$FusionPair))
ctrp_fus$G2P_Drug_FusionPair <- ifelse(ctrp_fus$Drug_FusionPair %in% unique(as.character(g2p_fus$Drug_FusionPair)), "Yes", "No")
ctrp_fus$G2P_Fusion <- ifelse(ctrp_fus$FusionPair %in% unique(as.character(g2p_fus$FusionPair)), "Yes", "No")
ctrp_fus$Database <- "CTRP"

gdsc_fus <- auc %>% filter(Database == "GDSC") %>% select(Cell_line, CCLE_histology, TCGA_classification, CCLE_name, DepMap_ID, Drug, AUC_IC50)
gdsc_fus_grid <- expand.grid("DepMap_ID" = unique(gdsc_fus$DepMap_ID), "FusionPair" = unique(fusions_filt$FusionPair))
gdsc_fus <- merge(merge(gdsc_fus, gdsc_fus_grid, by = "DepMap_ID", all = TRUE), fusions_filt, by = c("DepMap_ID", "CCLE_name", "CCLE_histology", "Cell_line", "FusionPair"), all = TRUE)
gdsc_fus$Fusion_Status <- ifelse(!is.na(gdsc_fus$LeftGene), "Yes", "No")
gdsc_fus$Drug_FusionPair <- ifelse(is.na(gdsc_fus$Drug) | is.na(gdsc_fus$FusionPair), NA, paste0(gdsc_fus$Drug, "_", gdsc_fus$FusionPair))
gdsc_fus$G2P_Drug_FusionPair <- ifelse(gdsc_fus$Drug_FusionPair %in% unique(as.character(g2p_fus$Drug_FusionPair)), "Yes", "No")
gdsc_fus$G2P_Fusion <- ifelse(gdsc_fus$FusionPair %in% unique(as.character(g2p_fus$FusionPair)), "Yes", "No")
gdsc_fus$Database <- "GDSC"

@@

```{r}
fus <- readRDS("./data_munging/rds/fusions_merged.rds") %>% drop_na(Drug)
fus_summary <- fus %>% count(Database, Drug, FusionPair, Fusion_Status)
```

@@

crispr <- MergeOmicsG2P(database = "CRISPR", maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = g2p_genes, outfile = "./data_munging/19Q2/crispr_data_19Q2.rds")
write.csv.gz(crispr %>% drop_na(Score), "./../htscreens_giant_files/crispr_data_19Q2.csv.gz")
```

@@

```{r wilcoxon_heatmap_function_EDIT, eval=FALSE}
MakeWilcoxonHeatmap <- function(signif_df, feature, row_names, col_title, color, legend_title, legend_breaks, row_title_txt, radius) {
  if(!is.na(str_match(string = tolower(feature), pattern = "mut"))) {
    signif_grid <- dcast(signif_df, Gene ~ Drug, value.var = "p_log10", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
    signif_grid <- signif_grid[g2p_gene_order, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid <- signif_grid[complete.cases(signif_grid), ]
    signif_grid[signif_grid < -log10(0.05)] <- NA

    signif_grid_annot <- dcast(signif_df, Gene ~ Drug, value.var = "Direction_Curated", fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
    signif_grid_annot <- signif_grid_annot[g2p_gene_order, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid_annot <- signif_grid_annot[complete.cases(signif_grid_annot), ]
    # signif_grid_annot[signif_grid_annot == "No"] <- ""
    # signif_grid_annot[signif_grid_annot == "Yes"] <- "O"
  }
  if(!is.na(str_match(string = tolower(feature), pattern = "fus"))) {
    signif_grid <- dcast(signif_df, FusionPair ~ Drug, value.var = "p_log10", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-FusionPair)
    signif_grid <- signif_grid[, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid <- signif_grid[complete.cases(signif_grid), ]
    signif_grid[signif_grid < -log10(0.05)] <- NA

>>>>

# Manually assign biomarker_types based on a combination of biomarker_type derived from the JSON string and patterns in the Feature.Names
biomarker_type_curated <- 
  ifelse(!is.na(str_match(string = tolower(g2p$biomarker_type), pattern = "copy number variant|amplification")[,1]) | !is.na(str_match(string = tolower(g2p$Feature.Names), pattern = "copy number variant|amplification")), "Amplification/CNV",
  ifelse(!is.na(str_match(string = toupper(g2p$Feature.Names), pattern = "[A-Z0-9|A-Z]+-[A-Z0-9|A-Z]+|FUSION")[,1]) & is.na(str_match(string = g2p$Feature.Names, pattern = "[0-9]+-[0-9]+")) | g2p$biomarker_type == "fusion" | g2p$biomarker_type == "rearrange", "Fusion",
  ifelse(!is.na(str_match(string = g2p$Feature.Names, pattern = "[A-Z][0-9]+([A-Z]|[*])|[0-9A-Z]+ [A-Z][0-9]+|[A-Z][0-9]+$")[,1]), "Mutation",
  ifelse(!is.na(str_match(string = tolower(g2p$Feature.Names), pattern = "expression|methylation")[,1]), "Expression",
  ifelse(!is.na(str_match(string = g2p$biomarker_type, pattern = "intron|insertion|deletion|mutant")[,1]), "Mutation",
  ifelse(!is.na(str_match(string = tolower(g2p$Feature.Names), pattern = "codon|positive|biallelic inactivation|del|ins|dup|mut|frameshift|missense|variant|[0-9]+fs|>")[,1]), "Mutation",
  ifelse(!is.na(str_match(string = g2p$Feature.Names, pattern = "^KIT$|^BRCA$")[,1]), "Mutation", NA)))))))
g2p$biomarker_type_curated <- biomarker_type_curated[,1]

# Edit direction column
g2p <- add_column(g2p, Direction_Curated = NA, .after = "Direction")
g2p$Direction_Curated <- ifelse(!is.na(str_match(string = tolower(g2p$Direction), pattern = "resist")[,1]), "resistance", ifelse(!is.na(str_match(string = tolower(g2p$Direction), pattern = "sensitive|sensitivity|responsive")[,1]), "sensitivity", "none"))

@@

Mutation calls (coding region, germline filtered):
```{r mutation_status, eval=FALSE}
maf_raw <- read.delim("./data_munging/19Q2/CCLE_mutations_20190514.csv.gz", header = TRUE, sep = ",")
maf_raw[, 1:2] <- NULL
colnames(maf_raw)[colnames(maf_raw) == "Hugo_Symbol"] <- "Gene"
maf_raw <- unfactorize(maf_raw)

# Select columns for maf_changes_df
maf_changes <- select(maf_raw, Gene, DepMap_ID, Protein_Change) %>% drop_na(Protein_Change) %>% filter(Protein_Change != "" & Gene %in% g2p_genes)
maf_changes <- unique(maf_changes)
maf_changes$Protein_Change <- gsub(pattern = "^p.", replacement = "", x = maf_changes$Protein_Change)
maf_changes$Keep <- ifelse(!is.na(str_match(string = maf_changes$Protein_Change, pattern = "[A-Z][0-9]+([A-Z]|[*])")[,1]), "Yes", "No")
maf_changes_df <- filter(maf_changes, Keep == "Yes")
maf_changes_df$Keep <- NULL
maf_changes_df$Gene_Protein_Change <- with(maf_changes_df, paste0(Gene, ":", Protein_Change))
saveRDS(maf_changes_df, "./data_munging/19Q2/maf_changes_df_19Q2.rds", compress = "xz")

# Select columns for maf_df
maf <- subset(maf_raw, select = c("Gene", "DepMap_ID"))
maf$Mutation <- 1
maf_grid <- expand.grid("Gene" = unique(maf$Gene), "DepMap_ID" = unique(maf$DepMap_ID))
maf_df <- merge(maf, maf_grid, by = c("Gene", "DepMap_ID"), all = TRUE)
maf_df[is.na(maf_df)] <- 0
saveRDS(maf_df, "./data_munging/19Q2/maf_df_19Q2.rds", compress = "xz")
```

@@

```{r omics_functions}
MergeOmicsG2P <- function(database, dataframe, maf_df, ge_df, cn_df, gene_list, outfile) {
  # Merge mutation status, CN, and GE omics data with drugscreen results and metadata
  if(database %in% c("CCLE", "CTRP", "GDSC")) {
    print(paste0("Processing ", database, " data."))

    grid <- expand.grid("Cell_line" = unique(dataframe$Cell_line), "Gene" = gene_list)
    data <-  merge(dataframe, grid, by = "Cell_line", all.y = TRUE)
    print("Merged grid.")
    # Merge metadata
    data <-  merge(data, select(ccl_meta, Cell_line, CCLE_histology, CCLE_name, DepMap_ID, disease), by = c("Cell_line", "CCLE_histology", "CCLE_name", "DepMap_ID"), all.y = TRUE)
    ## Add columns for G2P drug-gene and drug-gene-lineage sorting
    data$Drug_Gene <- paste0(data$Drug, "_", data$Gene)
    data$G2P_Drug_Gene <- ifelse(data$Drug_Gene %in% g2p_dg, "Yes", "No")
    data$Drug_Gene_Lin <- paste0(data$Drug, "_", data$Gene, "_", data$disease)
    data$G2P_Drug_Gene_Lin <- ifelse(data$Drug_Gene_Lin %in% g2p_dgl, "Yes", "No")
    data$Gene_Lin <- paste0(data$Gene, "_", data$Gene, "_", data$disease)
    data$G2P_Gene_Lin <- ifelse(data$Gene_Lin %in% g2p_gl, "Yes", "No")
    print("Merged metadata.")
    # Merge mutation calls
    data <- unfactorize(data)
    data <- merge(data, maf_df, by = c("Gene", "DepMap_ID"), all.x = TRUE)
    data$Mutation <- ifelse(is.na(data$Mutation), 0, data$Mutation)
  }
  if(database == "CRISPR") {
    print(paste0("Processing ", database, " data."))
    # Melt CRISPR dataset for merging
    data <- melt(dataframe, id.vars = "DepMap_ID", measure.vars = colnames(crispr_all)[2:ncol(crispr_all)], variable.name = "Gene", value.name = "Score")
    # Merge cell line metadata
    data <- merge(data, select(ccl_meta, Cell_line, CCLE_histology, CCLE_name, DepMap_ID, disease), by = "DepMap_ID", all.y = TRUE)
    print("Merged metadata.")
    # Merge mutation calls
    data <- unfactorize(data)
    data <- merge(data, maf_df, by = c("Gene", "DepMap_ID"), all.x = TRUE)
    data$Mutation <- ifelse(is.na(data$Mutation), 0, data$Mutation)
  }

@@

```{r adjust_drugscreen_data, eval=FALSE}
ccle_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CCLE_drug_response_metrics.csv.gz", sep = ",")
ctrp_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/CTRP_drug_response_metrics.csv.gz", sep = ",")
gdsc_all <- read.delim("./../htscreens_giant_files/pozdeyev_data/GDSC_drug_response_metrics.csv.gz", sep = ",")

auc <- bind_rows(ccle_all, ctrp_all, gdsc_all) %>% drop_na(Drug_name)
auc$Drug <- tolower(auc$Drug_name)
auc$Cell_line <- toupper(gsub("-", "", auc$Cell_line))
auc$Cell_line <- ifelse(auc$Cell_line == "K052", "KO52", auc$Cell_line)
auc$Cell_line <- ifelse(auc$Cell_line == "OVCAR3", "NIHOVCAR3", auc$Cell_line)
auc$Cell_line <- ifelse(auc$Cell_line == "OVCAR3", "NIHOVCAR3", auc$Cell_line)
auc$Cell_line <- ifelse(auc$Cell_line == "TI73", "T173", auc$Cell_line)
auc <- merge(ccl_histologies, auc, by = "Cell_line", all.y = TRUE)
colnames(auc)[colnames(auc) == "AUC_IC50"] <- "Score"
saveRDS(auc, "./../htscreens_giant_files/pozdeyev_data/pozdeyev2016_auc_all.rds", compress = "xz")

ccle_all <- auc %>% filter(Database == "CCLE") %>% select(Cell_line, CCLE_histology, TCGA_classification, CCLE_name, DepMap_ID, Drug, Score)
saveRDS(ccle_all, "./../htscreens_giant_files/pozdeyev_data/pozdeyev2016_ccle_all_edited.rds", compress = "xz")
ctrp_all <- auc %>% filter(Database == "CTRP") %>% select(Cell_line, CCLE_histology, TCGA_classification, CCLE_name, DepMap_ID, Drug, Score)
saveRDS(ctrp_all, "./../htscreens_giant_files/pozdeyev_data/pozdeyev2016_ctrp_all_edited.rds", compress = "xz")
gdsc_all <- auc %>% filter(Database == "GDSC") %>% select(Cell_line, CCLE_histology, TCGA_classification, CCLE_name, DepMap_ID, Drug, Score)
saveRDS(gdsc_all, "./../htscreens_giant_files/pozdeyev_data/pozdeyev2016_gdsc_all_edited.rds", compress = "xz")
```

`r nrow(filter(test, n != 1))`/`r nrow(test)` gene-cell line combinations have more than one mutation in the MAF file. This is corrected in the point mutation filtering done below.
```{r merge_drug_data, eval=FALSE}
ccle_all <- readRDS("./../htscreens_giant_files/pozdeyev_data/pozdeyev2016_ccle_all_edited.rds")
ctrp_all <- readRDS("./../htscreens_giant_files/pozdeyev_data/pozdeyev2016_ctrp_all_edited.rds")
gdsc_all <- readRDS("./../htscreens_giant_files/pozdeyev_data/pozdeyev2016_gdsc_all_edited.rds")

ccle <- MergeOmicsG2P(database = "CCLE", dataframe = ccle_all, maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = g2p_genes, outfile = "./data_munging/19Q2/ccle_data_19Q2.rds")
write.csv.gz(ccle %>% drop_na(Score), "./data_munging/19Q2/ccle_data_19Q2.csv.gz")

ctrp <- MergeOmicsG2P(database = "CTRP", dataframe = ctrp_all, maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = g2p_genes, outfile = "./../htscreens_giant_files/ctrp_data_19Q2.rds")
write.csv.gz(ctrp %>% drop_na(Score), "./../htscreens_giant_files/ctrp_data_19Q2.csv.gz")

gdsc <- MergeOmicsG2P(database = "GDSC", dataframe = gdsc_all, maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = g2p_genes, outfile = "./data_munging/19Q2/gdsc_data_19Q2.rds")
write.csv.gz(gdsc %>% drop_na(Score), "./../htscreens_giant_files/gdsc_data_19Q2.csv.gz")

# all_genes <- intersect(unique(maf_df$Gene), intersect(unique(cn_melt$Gene), unique(ge_melt$Gene)))
# ccle <- MergeOmicsG2P(database = "CCLE", maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = all_genes, outfile = "./../htscreens_giant_files/ccle_data_allgenes_19Q2.rds")
# ctrp <- MergeOmicsG2P(database = "CTRP", maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = all_genes, outfile = "./../htscreens_giant_files/ctrp_data_allgenes_19Q2.rds")
# gdsc <- MergeOmicsG2P(database = "GDSC", maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = all_genes, outfile = "./../htscreens_giant_files/gdsc_data_allgenes_19Q2.rds")
```

@@

Merge fusions:
```{r merge_fusions, eval=FALSE}
ccle_all <- readRDS("./../htscreens_giant_files/pozdeyev_data/pozdeyev2016_ccle_all_edited.rds")
ctrp_all <- readRDS("./../htscreens_giant_files/pozdeyev_data/pozdeyev2016_ctrp_all_edited.rds")
gdsc_all <- readRDS("./../htscreens_giant_files/pozdeyev_data/pozdeyev2016_gdsc_all_edited.rds")
fusions_filt <- filter(fusions, FusionPair %in% unique(as.character(g2p_fus$FusionPair)))

ccle_fus_grid <- expand.grid("DepMap_ID" = unique(ccle_all$DepMap_ID), "FusionPair" = unique(fusions_filt$FusionPair))
ccle_fus <- merge(merge(ccle_all, ccle_fus_grid, by = "DepMap_ID", all = TRUE), fusions_filt, by = c("DepMap_ID", "CCLE_name", "CCLE_histology", "Cell_line", "FusionPair"), all = TRUE)
ccle_fus$Fusion_Status <- ifelse(!is.na(ccle_fus$LeftGene), "Yes", "No")
ccle_fus$Drug_FusionPair <- ifelse(is.na(ccle_fus$Drug) | is.na(ccle_fus$FusionPair), NA, paste0(ccle_fus$Drug, "_", ccle_fus$FusionPair))
ccle_fus$G2P_Drug_FusionPair <- ifelse(ccle_fus$Drug_FusionPair %in% unique(as.character(g2p_fus$Drug_FusionPair)), "Yes", "No")
ccle_fus$G2P_Fusion <- ifelse(ccle_fus$FusionPair %in% unique(as.character(g2p_fus$FusionPair)), "Yes", "No")
ccle_fus$Database <- "CCLE"

ctrp_fus_grid <- expand.grid("DepMap_ID" = unique(ctrp_all$DepMap_ID), "FusionPair" = unique(fusions_filt$FusionPair))
ctrp_fus <- merge(merge(ctrp_all, ctrp_fus_grid, by = "DepMap_ID", all = TRUE), fusions_filt, by = c("DepMap_ID", "CCLE_name", "CCLE_histology", "Cell_line", "FusionPair"), all = TRUE)
ctrp_fus$Fusion_Status <- ifelse(!is.na(ctrp_fus$LeftGene), "Yes", "No")
ctrp_fus$Drug_FusionPair <- ifelse(is.na(ctrp_fus$Drug) | is.na(ctrp_fus$FusionPair), NA, paste0(ctrp_fus$Drug, "_", ctrp_fus$FusionPair))
ctrp_fus$G2P_Drug_FusionPair <- ifelse(ctrp_fus$Drug_FusionPair %in% unique(as.character(g2p_fus$Drug_FusionPair)), "Yes", "No")
ctrp_fus$G2P_Fusion <- ifelse(ctrp_fus$FusionPair %in% unique(as.character(g2p_fus$FusionPair)), "Yes", "No")
ctrp_fus$Database <- "CTRP"

gdsc_fus_grid <- expand.grid("DepMap_ID" = unique(gdsc_all$DepMap_ID), "FusionPair" = unique(fusions_filt$FusionPair))
gdsc_fus <- merge(merge(gdsc_all, gdsc_fus_grid, by = "DepMap_ID", all = TRUE), fusions_filt, by = c("DepMap_ID", "CCLE_name", "CCLE_histology", "Cell_line", "FusionPair"), all = TRUE)
gdsc_fus$Fusion_Status <- ifelse(!is.na(gdsc_fus$LeftGene), "Yes", "No")
gdsc_fus$Drug_FusionPair <- ifelse(is.na(gdsc_fus$Drug) | is.na(gdsc_fus$FusionPair), NA, paste0(gdsc_fus$Drug, "_", gdsc_fus$FusionPair))
gdsc_fus$G2P_Drug_FusionPair <- ifelse(gdsc_fus$Drug_FusionPair %in% unique(as.character(g2p_fus$Drug_FusionPair)), "Yes", "No")
gdsc_fus$G2P_Fusion <- ifelse(gdsc_fus$FusionPair %in% unique(as.character(g2p_fus$FusionPair)), "Yes", "No")
gdsc_fus$Database <- "GDSC"

@@

```{r load_fusion_data}
fus <- readRDS("./data_munging/rds/fusions_merged.rds") %>% drop_na(Drug)
fus_summary <- fus %>% count(Database, Drug, FusionPair, Fusion_Status)

```

Merge protein changes: **NEED TO MAKE G2P_prot DATAFRAME (6/4/19)**
```{r merge_protein_changes, eval=FALSE}
ccle_all <- readRDS("./../htscreens_giant_files/pozdeyev_data/pozdeyev2016_ccle_all_edited.rds")
ctrp_all <- readRDS("./../htscreens_giant_files/pozdeyev_data/pozdeyev2016_ctrp_all_edited.rds")
gdsc_all <- readRDS("./../htscreens_giant_files/pozdeyev_data/pozdeyev2016_gdsc_all_edited.rds")
maf_changes_df <- readRDS("./data_munging/19Q2/maf_changes_df_19Q2.rds")

ccle_prot_grid <- expand.grid("DepMap_ID" = unique(ccle_all$DepMap_ID), "Protein_Change" = unique(maf_changes_df$Protein_Change))
ccle_prot <- merge(merge(ccle_all, ccle_prot_grid, by = "DepMap_ID", all = TRUE), maf_changes_df, by = c("DepMap_ID", "CCLE_name", "CCLE_histology", "Cell_line", "Protein_Change"), all = TRUE)
ccle_prot$Protein_Change_Status <- ifelse(!is.na(ccle_prot$Protein_Change), "Yes", "No")
ccle_prot$Drug_Protein_Change <- ifelse(is.na(ccle_prot$Drug) | is.na(ccle_prot$Protein_Change), NA, paste0(ccle_prot$Drug, "_", ccle_prot$Protein_Change))
# ccle_prot$G2P_Drug_Protein_Change <- ifelse(ccle_prot$Drug_Protein_Change %in% unique(as.character(g2p_prot$Drug_Protein_Change)), "Yes", "No")
# ccle_prot$G2P_Protein_Change <- ifelse(ccle_prot$Protein_Change %in% unique(as.character(g2p_prot$Protein_Change)), "Yes", "No")
ccle_prot$Database <- "CCLE"

```

@@

crispr <- MergeOmicsG2P(database = "CRISPR", dataframe = crispr_all, maf_df = maf_df, ge_df = ge_melt, cn_df = cn_melt, gene_list = g2p_genes, outfile = "./data_munging/19Q2/crispr_data_19Q2.rds")
write.csv.gz(crispr %>% drop_na(Score), "./../htscreens_giant_files/crispr_data_19Q2.csv.gz")
```

@@
